---
title: "Identifying representation sequences of trajectories between IMD deciles"
author: "F Rowe"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---


```{r}
library(tidyverse)
library(TraMineR) # this is used for the sequence analysis
library(cluster)
library(WeightedCluster)
library(viridis)
library(scales)
library(ggthemes)
library(showtext)
```
Set font style
```{r}
# load font
font_add_google("Roboto Condensed", "robotocondensed")
font_add_google("Fira Sans Extra Condensed","cond")
font_add_google("Fira Sans","fira")
# automatically use showtext to render text
showtext_auto()
```

## read the file containing the IMD deciles by LSOA

```{r}
IMD_deciles <- read_csv('/Users/franciscorowe/Dropbox/Francisco/Research/in_progress/left_behind_places/data/IMDTraj.csv')
head(IMD_deciles)
```

## Create a table with unique sequences and their frequency
```{r}
# Create a new column that keeps only the sequences that I create manually
IMD_deciles$sequence <- paste(IMD_deciles$"2010_DECILE", IMD_deciles$"2015_DECILE", IMD_deciles$"2019_DECILE", sep="-")

# count the unique sequences
db1 <- aggregate(data.frame(
  count = IMD_deciles$sequence),
  list(value = IMD_deciles$sequence),
  length)
head(db1)

```




```{r}
# Create a sequence objectby specifying the values and the weights
# specify also the colour of each decile
seq.cl <- seqdef(db1$value, weights = db1$count, cnames = c("2010", "2015", "2019"),
                 cpal =c("1" = "#7f3b08", 
                         "2" = "#b35806",
                         "3" = "#e08214",
                         "4" = "#fdb863",
                         "5" = "#fee0b6",
                         "6" = "#d8daeb",
                         "7" = "#b2abd2",
                         "8" = "#8073ac",
                         "9" = "#542788",
                         "10" = "#2d004b"))

```



```{r}
# Calculate transition rates
subs_costs <- seqsubm(seq.cl, method = "TRATE")
subs_costs <- round(subs_costs, 2)

# Calculate the unique distance matrix
seq.OM <- seqdist(seq.cl,
                  method = "OM",
                  sm = subs_costs)
```


## I run a few examples of different number of clusters and assess their performance
```{r}
# info on weighted clustering https://cran.r-project.org/web/packages/WeightedCluster/vignettes/WeightedCluster.pdf
# PAM clustering
for (i in 3:15) {
  pamwardclust <- wcKMedoids(seq.OM, k= i, weights=db1$count)
  print(paste("number of k =", i,": Average Silhouette Width (weighted) is" , pamwardclust$stats[5]))

}
```

Table 3 in the manual of weighted clustering package https://cran.r-project.org/web/packages/WeightedCluster/vignettes/WeightedCluster.pdf suggests that:

| ASW         | Interpretation proposed                   |
| ----------- | ----------------------------------------- |
| 0.71 − 1.00 | Strong structure identified               |
| 0.51 − 0.70 | Reasonable structure identified           |
| 0.26 − 0.50 | Structure is weak and could be artificial |
| ≤ 0.25      | Try other algorithms. No structure        |

I think the best clustering solution is  with k = 9
However I plot with k=6 and k=8 too


### with k = 6
```{r}
pamwardclust6 <- wcKMedoids(seq.OM, k = 6, weights=db1$count)

# give names to clusters
# 1 is the cluster with the most deprived areas in 2010 and 6 is the cluster with the least deprived areas in 2010
pamwardclust6$clustering <- factor(pamwardclust6$clustering, levels=c(1,
                                                                    38,
                                                                    69,
                                                                    127,
                                                                    209,
                                                                    30),
                                  labels=c("1", "2", "3", "4", "5", "6"))

# calculate silouette score to use later in graphs
sil <- wcSilhouetteObs(seq.OM, pamwardclust6$clustering, weights=db1$count, measure="ASWw")
seqIplot(seq.cl, group = pamwardclust6$clustering, border =NA, cex.axis=1.5, cex.lab=1.5)
```

### with k = 8
```{r}
pamwardclust8 <- wcKMedoids(seq.OM, k = 8, weights=db1$count)
# give names to clusters
# 1 is the cluster with the most deprived areas in 2010 and 8 is the cluster with the least deprived areas in 2010
pamwardclust8$clustering <- factor(pamwardclust8$clustering, levels=c(1,
                                                                    38,
                                                                    69,
                                                                    98,
                                                                    159,
                                                                    209,
                                                                    267,
                                                                    18),
                                  labels=c("1", "2", "3", "4", "5", "6", "7", "8"))


# calculate silouette score to use later in graphs
sil <- wcSilhouetteObs(seq.OM, 
                       pamwardclust8$clustering, 
                       weights=db1$count, 
                       measure="ASWw")
```


```{r}
for (i in unique(pamwardclust8$clustering)) {
  seqIplot(seq.cl[pamwardclust8$clustering == i,],
           border =NA,
           with.legend = FALSE,
           sortv=sil, cex.axis=1.5,
           cex.lab=1.5)
} 
  
seqlegend(seq.cl, ncol = 5, cex = 1.2)
```

### with k = 9
```{r}
pamwardclust9 <- wcKMedoids(seq.OM, 
                            k = 9, 
                            weights=db1$count)

# give names to clusters
# 1 is the cluster with the most deprived areas in 2010 and 8 is the cluster with the least deprived areas in 2010
pamwardclust9$clustering <- factor(pamwardclust9$clustering, levels=c(1,
                                                                    38,
                                                                    69,
                                                                    98,
                                                                    136,
                                                                    197,
                                                                    240,
                                                                    267,
                                                                    18),
                                  labels=c("1", "2", "3", "4", "5", "6", "7", "8", "9"))

# calculate silouette score to use later in graphs
sil <- wcSilhouetteObs(seq.OM, 
                       pamwardclust9$clustering,
                       weights=db1$count, measure="ASWw")
```

```{r}
for (i in unique(pamwardclust9$clustering)) {
  seqIplot(seq.cl[pamwardclust9$clustering == i,], 
           border =NA, 
           with.legend = FALSE, 
           sortv=sil, 
           cex.axis=1.5, 
           cex.lab=1.5)
} 
  
seqlegend(seq.cl, ncol = 5, cex = 1.2)
```

### with k = 10
```{r}
pamwardclust10 <- wcKMedoids(seq.OM, 
                            k = 10, 
                            weights=db1$count)

# give names to clusters
# 1 is the cluster with the most deprived areas in 2010 and 8 is the cluster with the least deprived areas in 2010
pamwardclust10$clustering <- factor(pamwardclust10$clustering, levels=c(1,
                                                                    38,
                                                                    69,
                                                                    98,
                                                                    131,
                                                                    164,
                                                                    203,
                                                                    240,
                                                                    267,
                                                                    18),
                                  labels=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"))

# calculate silouette score to use later in graphs
sil <- wcSilhouetteObs(seq.OM, 
                       pamwardclust10$clustering,
                       weights=db1$count, measure="ASWw")
```


```{r}
mypath <- file.path("../outcomes/",
                      paste("seqiplot_6c.jpg", sep = ""))
  
  jpeg(file=mypath, width = 1000, height = 1000)
  seqIplot(seq.cl, 
           group = pamwardclust6$clustering, 
           border =NA, 
           cex.axis=1.5, 
           cex.lab=1.5)
  
  dev.off()
```
```{r}
mypath <- file.path("../outcomes/",
                      paste("seqiplot_8c.jpg", sep = ""))
  
  jpeg(file=mypath, width = 1000, height = 1000)
  seqIplot(seq.cl, 
           group = pamwardclust8$clustering, 
           border =NA, 
           cex.axis=1.5, 
           cex.lab=1.5)
  
  dev.off()
```

```{r}
mypath <- file.path("../outcomes/",
                      paste("seqiplot_9c.jpg", sep = ""))
  
  jpeg(file=mypath, width = 1000, height = 1000)
  seqIplot(seq.cl, 
           group = pamwardclust9$clustering, 
           border =NA,
           sortv=sil,
           cex.axis=1.5, 
           cex.lab=1.5)
  
  dev.off()
```
```{r}
mypath <- file.path("../outcomes/",
                      paste("seqiplot_10c.jpg", sep = ""))
  
  jpeg(file=mypath, width = 1000, height = 1000)
  seqIplot(seq.cl, 
           group = pamwardclust10$clustering, 
           border =NA,
           sortv=sil,
           cex.axis=1.5, 
           cex.lab=1.5)
  
  dev.off()
```

```{r}
# Attach the cluster label next to each unique sequence
# there are three new columns corresponding to three cluster solutions (i.e. 6, 8 and 9)

db1$clst6 <- pamwardclust6$clustering
db1$clst8 <- pamwardclust8$clustering
db1$clst9 <- pamwardclust9$clustering
db1$clst10 <- pamwardclust10$clustering

```


```{r}
# Create a dataframe that includes all sequences with their corresponding cluster number
IMD_deciles_clst <- merge(x = IMD_deciles, 
                          y = db1, 
                          by.x = "sequence",
                          by.y = "value", 
                          all.x = TRUE)

head(IMD_deciles_clst)

write.csv(IMD_deciles_clst, "../data/imd_clstid.csv")

```



```{r}
elbow_plot <- fviz_nbclust(seq.OM, 
                           cluster::pam, 
                           method = "wss",
                           k.max = 20)


df_elbow <- do.call(data.frame, elbow_plot$data)
df_elbow <- df_elbow %>% 
  mutate(method = "wss",
         value = y / 100000) %>% 
  dplyr::select( c(clusters, value, method))

print(elbow_plot)
```
```{r}
silhouette_plot <- fviz_nbclust(seq.OM, 
                                cluster::pam, 
                                method = "silhouette",
                                k.max = 20)

df_silhouette <-do.call(data.frame, silhouette_plot$data)

df_silhouette <- df_silhouette %>% 
  mutate(method = "silhouette") %>% 
  rename(value = y)
  
print(silhouette_plot)
```

```{r}
gap_plot <- fviz_nbclust(seq.OM, 
                         cluster::pam, 
                         method = "gap_stat",
                         k.max = 20)

df_gap <- gap_plot$data

df_gap <- df_gap %>% 
  dplyr::select(
  c(clusters, gap)
) %>% mutate(
  method = "gap"
) %>% 
  rename(value = gap)
  
print(gap_plot)
```
```{r}
df_scores <- rbind(df_elbow, df_silhouette, df_gap)
df_scores$clusters <- as.integer(as.character(df_scores$clusters))
```

```{r}
# Create the line plot
ggplot(df_scores, 
       aes(x = clusters, y = value, color = method)) +
  geom_line(size = 2) +
  labs(x = "Number of clusters k", y = "Score") +
  scale_color_manual(values = c("wss" = "#E69F00", "silhouette" = "#009E73", "gap" = "#0072B2")) +
  theme_tufte() + 
  theme(
    text = element_text(family = "robotocondensed"),
    axis.text.y = element_text(size = 20),
    axis.title.y = element_text(size = 22),
    axis.text.x = element_text(size = 20),
    axis.title.x = element_text(size = 22),
    legend.position = "bottom",
    legend.text = element_text(size = 22),
    legend.title=element_blank()
    )

png("/Users/franciscorowe/Dropbox/Francisco/Research/in_progress/left_behind_places/outcomes/optimal_cluster.png",units="in", width=10, height=10, res=300)
last_plot()
dev.off()

```


```{r}
df

ggplot(df, aes(x = X, y = Y, color = Group)) +
  geom_line() +
  labs(x = "X Axis", y = "Y Axis", title = "Line Plot for Three Groups") +
  scale_color_manual(values = c("Group1" = "blue", "Group2" = "red", "Group3" = "green")) +
  theme_minimal()
```




